|

### 2021-09-28 03-Chp-Systematic-Review.Rmd
End of day:

There were 127 primary studies <!-- TODO check this -->that contained information on this relationship. I found that statins reduce the risk of Alzheimer's disease, but had no effect of vascular dementia. Lipids were not associated with any outcome. The methods used in some of the primary studies meant that I was less confident in the accuracy of their results.

@@

<!-- Where both the outcome and the exposure were dichotmous, or the -->

@@

@@

In response to this criticism, I develope a new visualisation tool was designed to allow for "paired" forest plots, as recommended by the ROB2 publication, where the risk-of-bias assessment is presented alongside the results.[@sterne2019] This tool was developed as an adjunct to this thesis to aid in creating standardised risk-of-bias figures,[@mcguinness2020robvisPaper] and the "paired" forest plot functionality grew out of a collaboration with other researchers to design a modular method for creating custom forest plots.[@zotero-14999] <!-- Apparently collaboration is a big thing they want to see --> A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this Chapter were created using this tool.

@@

@@

Additionally, I followedpreprints up over time to investigate whether all identified preprints included in the review were subsequently published (in which case preprints provide a snapshot into the future, and a systematic review update would capture these reports) or alternatively, if some preprints were not published, then preprints provide a distinct evidence source.

@@

<!--- TODO Cross check this with the actual numbers included --->

@@

(ref:prisma-flow-cap) PRISMA flow diagram illustrating how records moved through the systematic review process. The different contributions of standarad bibliographic databases and preprint servers to the review are indicated.

@@

Following full-text screening, __XXXX__ studies met the criteria for inclusion in the review.

Table \@ref(tab:studyCharacteristics-table) presents a summary of the characteristics of each study.

The majority of studies were non-randomised, with only two included randomised controlled trials, both of which examined all-cause dementia.

Similarly, the majority of the few included  and a relative small number of Mendelian randomisation studies. Of the non-randomised studies, many were assessments of the relationship with statis use.

The vast majority of studies examined either all-cause dementia or Alzheimer's disease, with only a small proportion examining vascular dementia (n=XXX, %).

@@

Of the studies examining lipid levels, by far the most common fraction assessed was total cholesterol. The association with triglyceride levels was not consider in many papers.
<!-- TODO Check this is actually true-->

Many studies using electronic health records as their data source did not accurately report the diagnostic criteria used to identify cases.

Midlife cholesterol levels were of particular interest, with several studies examining this age-range.

Three included reports were preprints (denoted in the Table \@ref(tab:studyCharacteristics-table) using an asterik), one of which had subsequently been published and was captured by the primary literature serachr. All three included preprints were obtained from the bioRxiv preprint server and described Mendelian randomisation analyses.

@@

-   PRISMA Flowchart (done)

-   Summary of types of study (half done)

-   Summary of locations (half done)

-   Summary of diagnostic criteria used (done)

@@

<!-- TODO Need to add asterik denoting preprints to table, and add a column on whether codes repoted for EHR -->

\blandscape{}
<!----------------------------------------------------------------------------->
(ref:studyCharacteristics-caption) Characteristics of included studies, stratified by study design.

@@

As illustrated in Figure \@ref(fig:cohortLocations), the majority of study cohorts were based in the developed world.

<!----------------------------------------------------------------------------->
```{r cohortLocationsSetup, include = FALSE}

```

@@

### Risk of bias {#risk-of-bias-subheading}

@@

### Statins

```{r primaryFigures, include = FALSE, message=FALSE,warning=FALSE}
```

<!-- Straight-forward random effects meta-analysis of binary exposures, using study defined measures of each. -->

In the two included RCTs, statin use was not associated with dementia outcomes

<!----------------------------------------------------------------------------->
(ref:statinsRCT-cap) __Random effects meta-analysis of statins on all-cause dementia:__

@@

```{r statinsRCT, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:statinsRCT-cap)', out.width='100%', fig.scap='(ref:statinsRCT-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forest_statins_rct.png"))
```
<!----------------------------------------------------------------------------->

In NRSI, statin use was associated with a reduce risk of all-cause dementia (`r estimate(0.76,0.66,0.88)`, Figure \@ref(fig:statinsNRSI)).

(ref:statinsNRSI-cap) __Forest plot showing effect of statins:__

(ref:statinsNRSI-scap) Forest plot showing effect of statins

```{r statinsNRSI, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:statinsNRSI-cap)', out.width='100%', fig.scap='(ref:statinsNRSI-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

In Mendelian randomisation studies, a similar blah blah blah

(ref:statinsMR-cap) __Forest plot showing effect of statins:__

(ref:statinsMR-scap) Forest plot showing effect of statins

```{r statinsMR, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:statinsMR-cap)', out.width='100%', fig.scap='(ref:statinsMR-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

@@

### Hypercholesterolemia and lipid levels

(ref:hypercholNRSE-cap) __Forest plot showing effect of statins:__

(ref:hypercholNRSE-scap) Forest plot showing effect of statins

```{r hypercholNRSE, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:hypercholNRSE-cap)', out.width='100%', fig.scap='(ref:hypercholNRSE-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

(ref:lipidsNRSE-cap) __Forest plot showing effect of statins:__

(ref:lipidsNRSE-scap) Forest plot showing effect of statins

```{r lipidsNRSE, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:lipidsNRSE-cap)', out.width='100%', fig.scap='(ref:lipidsNRSE-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

@@

(ref:lipidsDoseResponse-cap) __Forest plot showing effect of statins:__

(ref:lipidsDoseResponse-scap) Forest plot showing effect of statins

```{r lipidsDoseResponse, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:lipidsDoseResponse-cap)', out.width='100%', fig.scap='(ref:lipidsDoseResponse-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

<!-- ### Sources of heterogeneity -->

Planned to investigate effect of sex and age at baseline.

Plus other such as education level and baseline cognitive scores, but data were either not reported for most studies or when reported, were too diverse to synthesize.

@@

### Publication bias {#sys-rev-pub-bias}

@@

&nbsp;
<!----------------------------------------------------------------------------->
(ref:funnelStatinsAny-cap) Funnel plot of results examining the relationship between statins and any dementia

(ref:funnelStatinsAny-scap) Funnel plot of results examining the relationship between statins and any dementia
```{r funnelStatinsAny, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:funnelStatinsAny-cap)', out.width='100%', fig.scap='(ref:funnelStatinsAny-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/funnel_statins_any.png"))
```
<!----------------------------------------------------------------------------->
&nbsp;

### Added evidental value of including preprints {#sys-rev-including-preprints}

@@

I had intended to assess the added evidental value of including preprints as a literautre source.

As show in Figure \@ref(prisma-flow-fig), the number of results returned by the preprint searching was not substantial (bioRxiv = 256, medRxiv = 0). <!-- TODO check numbers -->

No unique preprints were identified as part of this review, with the published version of the few/sole <!-- TODO check which --> relevant preprint(s) being captured by the main search.

<!-- QUESTION Can I explore the amount of information added by the preprints quantiatively using the weights in the meta-analysis -->

@@

<!-- TODO Check inclusion of Rockwood 2002:  -->

@@

There was some evidence of protective effect of statins on all-cause and Alzheimer's disease dementia when looking at solely at observational studies. This finding was not supported by evidence from the single available RCT, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings are a

In addition,  there were some concerns over the potential for estimates to be missing from the meta-analysis of observational studies not at random, given the preferential reporting of

@@

This review has presented a summary of the available evidence on the association between lipids, and treatments that modify lipids such as statins, and the subsequent risk of dementia.

The distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the long follow-up, necessary due to the long latent period of the condition, makes trials logistically challenging and financial expensive. Similarly, Mendelian randomisation is a relatively new study design, and so only appears in the literature in recent years, driven by the growth of dementia GWASs (Figure \@ref(fig:typeByYear)).

@@

A common theme across the evidence base was a lack of data on the association of vascular dementia. This is particularly interesting given that lipids and statins are primarily related to vascular disease. There is the potential that studies encountered similar difficulties in address the unexpected results observed in the CPRD analysis in Chapter \@ref(cprd-analysis-heading), likely due to confounding by indication, and so may suffer from the "file-drawer effect".[@rosenthal1979] For vascular dementia, few Mendelian randomisation studies examined this outcome, primarily because of the absence (until recently) of genome wide association studies, which are used in two sample summary Mendelian randomisation studies.

One item of particular interest is the attenuation of any effects observed by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the Apoe4 gene region. As covered in the discussion, ApoE4 genotype is the major risk factor for Alzheimer's disease, but is also involved in cholesterol transport.

@@

This review did not include the commonly cited PROSPER study, which examined the effect of pravastatin on CVD risk and is regularly held up as  providing no evidence for an effect of the statin on dementia outcomes. While widely cited and included in the Cochrane review on this topic,[@mcguinness2016] only reported the mean change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test, etc.) over follow-up rather than incident. While this is a useful indicator of general cognitive decline, it is not equivalent to a dementia diagnosis using a , as cognitive tests should feed into a broader diagnostic pathway (see section \@ref(diagnostic-criteria)). As such, this trial did not met the inclusion criteria for this review.

@@

Questions over missing results - evidence from one of the conference abstract analysis pairs that a non-significant results are being suppressed.[@yamada2009; @yamada2009a] Tie this in with

@@

### Previous reviews {#rev-previous-reviews}

@@

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018a] However, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies, and explore the heterogeneity of results across different levels of risk of bias levels. Some previous reviews did assess risk of bias, but used non-domain based assessment tools, such as the Newcastle-Ottowa scale.[@anstey2015; @poly2020]

However, despite these differences in methodology, the duplication of work across reviews (including this review) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch could have been employed. Known as an umbrella review, or review-of-reviews, these studies use other systematic reviews rather than primary studies as the unit of analysis. This approach would have enable more efficient identification of relevant primary studies, to which the methods which sets this review apart from other published reviews could have been applied.

<!-- ### Triangulation across evidence sources -->

<!-- TODO Have a look at this - potentially move to triangulation chapter -->

<!-- One key question for which multiple distinct sources of evidence were available were those looking at Laz -->

<!-- Consideration of the potential impact of the magnitude and direct of residual confounders/bias is not a major stretch from what is already happening in the assessment of the quality of evidence (GRADE) framework. Within GRADE, the overall quality of evidence can be upgraded when there is deemed to be unmeasured or residual confounding variables which reduce the. For example, if the propensity to treatment is related to comorbidity burden, but those on treatment still have better outcomes then those on control, it is likely that the true effect of the intervention is being underestimated. [@guyatt2011] -->

<!-- Without our framework and as part of the risk-of-bias assessments reported in \@ref(risk-of-bias-subheading), I attempted to records the direction of the bias, so that it could feed into the triangulation. -->

As highlighted in Section \@ref(diverse-sources-preprints), this review explicitly sought to synthesize evidence across different publication statuses (preprinted vs. published). Using the tool described in Chapter \@ref(sys-rev-tools-intro), two preprint serves related to health and biomedical sciences were search as part of this review (see Appendix \@ref(appendix-medrxivr-code) for the code used to search the repositories). There were several relevant preprints captured by the search. The added evidential value of including these preprints was highlighted in Section \ref(sys-rev-including-preprints).

@@

@@

<!----------------------------------------------------------------------------->
(ref:mrDuplication-cap) Summary of duplication of Mendelian randomisation studies which used similar data from the Global Lipid

@@

Recommend future large scale GWAS of other dementia outcomes, notably vascular dementia, or the use of exisitng GWASes to  <!-- TODO At least on vascular dementia GWAS exists - find it and cite it here. -->

@@

>>>>

There were XXX primary studies <!-- TODO check this -->that contained information on this relationship. I found that statins reduce the risk of Alzheimer's disease, but had no effect of vascular dementia. Lipids were not associated with any outcome. The methods used in some of the primary studies meant that I was less confident in the accuracy of their results.

@@

@@

Finally, I investigated the potential for small study effects, which may be caused by publication bias, both visually using funnel plots and statistically using Egger's regression test.  <!-- TODO CITATION NEEDED - cite Jonathon on interpreting small study effects, and whatever citation is needed for  -->

@@

In response to this criticism, I developed a new visualisation tool was designed to allow for "paired" forest plots, as recommended by the ROB2 publication, where the risk-of-bias assessment is presented alongside the results.[@sterne2019] This tool was developed as an adjunct to this thesis to aid in creating standardised risk-of-bias figures,[@mcguinness2020robvisPaper] and the "paired" forest plot functionality grew out of a collaboration with other researchers to design a modular method for creating custom forest plots.[@zotero-14999] <!-- Apparently collaboration is a big thing they want to see --> A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this Chapter were created using this tool.

@@

__[Note: Julian, I am particularly interested in your feedback on this section, and the corresponding results section (Section \@ref(sys-rev-including-preprints)), as I am not convinced on the language I am using]__

@@

Additionally, I followed preprints up over time to investigate whether all identified preprints included in the review were subsequently published (in which case preprints provide a snapshot into the future, and a systematic review update would capture these reports) or alternatively, if some preprints were not published, then preprints provide a distinct evidence source.

@@

<!-- TODO Cross check this with the actual numbers included -->

@@

(ref:prisma-flow-cap) PRISMA flow diagram illustrating how records moved through the systematic review process. The different contributions of standard bibliographic databases and preprint servers to the review are indicated.

@@

Following full-text screening, 87 reports describing 78 unique studies met the criteria for inclusion in the review.[@ancelin2012a; @ancelin2013; @andrews2019a; @arvanitakis2008; @batty2014; @benn2017b; @bettermann2012a; @beydoun2011; @bruce2017; @burgess2017; @chao2015a; @chen2014; @chiang2007; @chitnis2015; @chou2014a; @chuang2015; @cramer2008; @dodge2011; @forti2010; @gnjidic2016; @gottesman2017; @gustafson2012; @haag2009; @hayden2006; @heartprotectionstudycollaborativegroup2002; @hendrie2015; @hippisley-cox2010a; @jick2000; @kimm2011; @kivipelto2001; @kivipelto2005a; @kuo2015; @li2004; @li2005a; @li2010; @liao2013; @liu2019a; @mainousa.g.2005; @mielke2005a; @mielke2010a; @mielke2011; @mukherjee2013; @muller2007; @noale2013; @notkola1998; @ostergaard2017; @pan2018; @parikh2011; @peters2009; @raffaitin2009; @rantanen2017; @rea2005a; @redelmeier2019; @reitz2004a; @reitz2010; @ridker2008; @ronnemaa2011; @schilling2017a; @smeeth2009a; @so2017; @solomon2007; @solomon2009a; @solomon2010; @sparks2008; @strand2013; @svensson2019; @szwast2007; @tan2003a; @tynkkynen2016; @tynkkynen2018; @wang2012; @whitmer2005a; @yang2015; @yoshitake1995; @zamrini2004; @zandi2005; @zhu2018a; @zimetbaum1992] Table \@ref(tab:studyCharacteristics-table) presents a summary of the characteristics of each study.

The majority of studies were non-randomised studies with the sole two included randomised controlled trials both examining all-cause dementia. Similarly, a relatively small number of Mendelian randomisation studies were identified, and several performed a two-sample MR analysis using the same summary statistics from published GWAS, leading to complications in synthesis (see Section\@ref()).

Of the non-randomised studies examining treatments that modify lipid levels, the overwhelming majority examined statin use. For exposue studies, total cholesterol and LDL-cw were the most frequently reported lipid elements.

In terms of outcomes, the vast majority of studies examined either all-cause dementia or Alzheimer's disease, with only a small proportion examining vascular dementia (n=XXX, %). Some rarer outcome classifications such as vascular-component or mixed dementia were also investigated.

@@

Many studies using electronic health records as their data source did not accurately report the diagnostic codes used to identify cases.

Midlife cholesterol levels were of particular interest, with several studies examining this age-range. <!-- TODO Talk a little bit more about age here -->

Three included reports were preprints (denoted in the Table \@ref(tab:studyCharacteristics-table) using an asterisk), one of which had subsequently been published and was captured by the primary literature search. All three included preprints were obtained from the bioRxiv preprint server and described Mendelian randomisation analyses.

@@

@@

<!-- TODO Need to add abbreviations to table -->

\blandscape{}
<!----------------------------------------------------------------------------->
(ref:studyCharacteristics-caption) Characteristics of included studies, stratified by study design. Note that three studies reported on multiple study designs, and these have been duplicated across the relevant sub-sections.

@@

As illustrated in Figure \@ref(fig:cohortLocations), the majority of reports described studies conducted in high-income countries.

<!----------------------------------------------------------------------------->
```{r cohortLocationsSetup, include = FALSE}
```

@@

### Risk of bias {#risk-of-bias-res}

@@

### All-cause dementia

__Statins__

```{r rctStatinDementia, include = FALSE, message=FALSE, warning=FALSE}
```

The two randomised controlled trials provided very weak evidence (`r rct_statin_acd`) of an effect on statin use on all-cause dementia risk (Figure \@ref(fig:rctStatinDementiaFig)).

(ref:statinsRCT-cap) Random effects meta-analysis of randomised controlled trials examining statin statins on all-cause dementia

@@

```{r rctStatinDementiaFig, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:statinsRCT-cap)', out.width='100%', fig.scap='(ref:statinsRCT-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_rct_statins_Dementia.png"))
```

```{r obsStatins, include = FALSE, message=FALSE, warning=FALSE}
```

In contrast, a meta-analysis of prospective observational studies provided evidence of a protective effect of statins use on all-cause dementia risk (`r obsStatins$Dementia`, Figure \@ref(fig:obsStatinDementiaFig)).

<!----------------------------------------------------------------------------->
(ref:obsStatinDementia-cap) Random effects meta-analysis of non-randomised studies examining the effect of statin use on all-cause dementia

(ref:obsStatinDementia-scap) Random effects meta-analysis of statins on all-cause dementia

```{r obsStatinDementiaFig, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:obsStatinDementia-cap)', out.width='100%', fig.scap='(ref:obsStatinDementia-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_obs_Statin-Ever_Dementia.png"))
```
<!----------------------------------------------------------------------------->

```{r mrStatins, include = FALSE, message=FALSE, warning=FALSE}
```

A single study examing the effect of HMGCR inhibition on any dementia, via a , and found weak evidence for an effect (`r mrStatin$Dementia`).

__Lipids__

* NRSE

* MR studies per 1-SD

In NRSI, statin use was associated with a reduce risk of all-cause dementia (`r obsStatins$Dementia`, Figure \@ref(fig:statinsNRSI)).

In Mendelian randomisation studies, a similar

@@

### Alzheimer's disease

There were no randomised trials of statin, or any other lipid regulating agent, use on Alzheimer's disease, though several observational studies reported on this outcome, providing evidence for a protective effect (`r obsStatins$AD`; Figure \@ref(fig:obsStatinADFig))

<!----------------------------------------------------------------------------->
(ref:obsStatinAD-cap) Random effects meta-analysis of non-randomised studies examining the effect of statin use on Alzheimer's disease

(ref:obsStatinAD-scap) Random effects meta-analysis of statins on Alzheimer's disease

```{r obsStatinADFig, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:obsStatinAD-cap)', out.width='100%', fig.scap='(ref:obsStatinAD-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_obs_Statin-Ever_AD.png"))
```
<!----------------------------------------------------------------------------->

Two Mendelian randomisation studies looked at specifically as a result of by HMGCR inhibition (mediated by a single SNP; rs17238484). The first used a one sample approach (SNP-exposure and SNP-outcome associations are estimated using the same dataset) in a large Copenhagen-based cohort, while the second made use of summary level data obtained from the Global Lipids Genetic Consortium (SNP-exposure) and the International Genomics of Alzheimer's Project (SNP-outcome).

<!----------------------------------------------------------------------------->
(ref:mrStatinDementia-cap) Random effects meta-analysis of LDL-c lowering via genetic HMGCR inhibition on Alzheimer's disease

(ref:mrStatinDementia-scap) Random effects meta-analysis of LDL-c lowering via genetic HMGCR inhibition on Alzheimer's disease

```{r mrStatinDementiaFig, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:mrStatinDementia-cap)', out.width='100%', fig.scap='(ref:mrStatinDementia-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_MR_HMGCR_AD.png"))
```
<!----------------------------------------------------------------------------->

__Lipids__

### Vascular dementia

__Statins__

<!----------------------------------------------------------------------------->
(ref:obsStatinVaD-cap) Random effects meta-analysis of non-randomised studies examining effect of statin use on vascular dementia

(ref:obsStatinVaD-scap) Random effects meta-analysis of statins on vascular dementia

```{r obsStatinDVaDFig, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:obsStatinVaD-cap)', out.width='100%', fig.scap='(ref:obsStatinDVaD-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_obs_Statin-Ever_VaD.png"))
```
<!----------------------------------------------------------------------------->

__Lipids__

@@

### Sources of heterogeneity

Data on another potentially important source of heterogeneity
Plus other such as education level and baseline cognitive scores, but data were either not reported for most studies or when reported, were too diverse to synthesize.

@@

### Sensitivity analysis

Given that the majority of the studies included in the review were of

### Small study effects {#sys-rev-pub-bias}

@@

<!----------------------------------------------------------------------------->
(ref:funnelStatinsAny-cap) Funnel plot of results examining the relationship between statins and any dementia

(ref:funnelStatinsAny-scap) Funnel plot of results examining the relationship between statins and any dementia

```{r funnelStatinsAny, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:funnelStatinsAny-cap)', out.width='100%', fig.scap='(ref:funnelStatinsAny-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_rct_statins_Dementia.png"))
```

### Added evidental value of including preprints {#sys-rev-including-preprints-res}

@@

As show in Figure \@ref(prisma-flow-fig), the number of hits returned by the preprint searching was not substantial (bioRxiv = 256, medRxiv = 0). From these hits, three preprints were included in the review as reports of eligible studies, of which two were unique reports not captured by the main search.

For the analysis of XX on YYY, preprints added ...

Investigation of the publication status of the two unique preprints indicated that one has since been published (in late 2019 date).<!-- TODO check publication date -->

<!-- QUESTION Can I explore the amount of information added by the preprints quantitatively using the weights in the meta-analysis -->

@@

<!-- TODO Check inclusion of Rockwood 2002:  -->
This review has presented a summary of the available evidence on the association between lipids, and treatments that modify lipids such as statins, and the subsequent risk of dementia.

The discussion seeks to  to summarise the key findings in terms of literature sources and results as reported. A detailed comparison across the evidence sources, exposure measures and sources of bias reported here is presented as part of the triangulation exercise (Chapter \@ref(tri-heading)).

@@

There was some evidence of protective effect of statins on all-cause and Alzheimer's disease dementia when looking at solely at observational studies. This finding was not supported by evidence from the two available RCTs, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings may be a result of heterogenity in exposure (e.g. mid-life in studies of lipids with late-life lipid reduction in RCTs) or alternatively due to biases within the non-randomised studies.

@@

The majority of studies were non-randomised studies of lipids, or treatments that affect lipid levels such as statins. This distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the long follow-up, necessary due to the long latent period of the condition, makes trials logistically challenging and financial expensive. Similarly, Mendelian randomisation is a comparatively new study design (as illustrated in Figure \@ref(fig:typeByYear)), and so only appears in the literature in recent years, driven by the availability of Alzheimer's disease summary genome wide association studies (GWAS) that form the basis of two-sample Mendelian randomisation approach.

@@

A common theme across the evidence base was a lack of data on the association of vascular dementia. This is particularly interesting given that lipids and statins are primarily related to vascular disease. __Can't look forward to other chapter__ There is the potential that studies encountered similar difficulties in address the unexpected results observed in the CPRD analysis in Chapter \@ref(cprd-analysis-heading), likely due to confounding by indication, and so may suffer from the "file-drawer effect".[@rosenthal1979] For vascular dementia, few Mendelian randomisation studies examined this outcome, primarily because of the absence (until recently) of GWAS of this outcome.

@@

Of note, this review did not include the commonly cited PROSPER study, which examined the effect of pravastatin on CVD risk,[@shepherd2002a] reporting on cognitive outcomes as one of several secondary outcomes. While widely cited in relation to the effect of statins on dementia and included in the Cochrane review of RCTs on this topic,[@mcguinness2016] the trial only reported on the change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test and others) over follow-up. Though an useful indicator of general cognitive decline, it is not equivalent to a dementia diagnosis using recognised criteria, as cognitive tests should feed into a broader diagnostic pathway (see Section \@ref(diagnostic-criteria)). As such, this trial did not met the inclusion criteria for this review.

@@

Questions over missing results - evidence from one of the conference abstract analysis pairs that a non-significant results are being suppressed.[@yamada2009; @yamada2009a] In addition, there were some concerns over the potential for estimates to be missing from the meta-analysis of observational studies not at random, given the preferential reporting of significant results observed in a number of analysis (see Section \@ref(risk-of-bias-res)).

@@

### Comparison with previous reviews {#rev-previous-reviews}

@@

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018a; @kuzma2018a] However, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies, and explore the heterogeneity of results across different levels of risk of bias levels. Some previous reviews did assess risk of bias, but used non-domain based assessment tools, such as the Newcastle-Ottowa scale.[@anstey2015; @poly2020]

Similar the on previous review of Mendelian randomisation studies examining risk factors for Alzhiemer's disease was conducted prior to the majority of MR studies included in this review being published, and extracted results including SNPs in APoE4 (see the following section for a discussion of the bias this introduces).

However, despite these differences in timescales and methodology, the duplication of work across reviews (including this review) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch could have been employed. Known as an umbrella review, or review-of-reviews, these studies use other systematic reviews rather than primary studies as the unit of analysis.[@aromataris2015; @smith2011] This approach would have enabled more efficient identification of relevant primary studies to which the methods which sets this review apart from other published reviews could have been applied.

@@

One item of particular interest is the attenuation of any effects observed by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the Apoe4 gene region. As covered in the introduction (see Section \@ref()), increasing number of ApoE4 alleles is a major independent risk factor for Alzheimer's disease, and so violates the exclusion restriction criteria of .

@@

<!----------------------------------------------------------------------------->
(ref:mrDuplication-cap) Summary of duplication of Mendelian randomisation studies which used summary statistics from the Global Lipid Genetics Consortium (GLGC) and the International Genomics of Alzheimer's Project (IGAP). Note that the Alzheimer’s Disease Genetics Consortium (ADGC) is one of the component consortia of IGAP.

@@

<!-- TODO Talk about discrepancies between results of different methods, but highlight that it is a problem as the information is being contributed by the same people. -->

Recommend future large scale GWAS of other dementia outcomes, notably vascular dementia, or the use of existing GWAS to  <!-- TODO At least on vascular dementia GWAS exists - find it and cite it here. -->

@@

As highlighted in Section \@ref(diverse-sources-preprints), this review explicitly sought to synthesize evidence across different publication statuses (preprinted vs. published). Using the tool described in Chapter \@ref(sys-rev-tools-intro), two preprint serves related to health and biomedical sciences were search as part of this review (see Appendix \@ref(appendix-medrxivr-code) for the code used to search the repositories). There were several relevant preprints captured by the search. The added evidential value of including these preprints was highlighted in Section \ref(sys-rev-including-preprints-res).

### 2021-09-28 06-Chp-Triangulation.Rmd
End of day:

@@

@@

@@

## Graphical representation

Where the direction of bias was unclear/could not be determined from the

A different approach was taken for the confounding versus other domains, as confounding will either pull to the left or right, regardless of where the effect estimate is, while other domains will pull towards/away from the null (e.g. non-differential misclassification). In this case, the program accounted for the position of the effect estimate when assigning a directional arrow to the

@@

@@

@@

Compare and contrast with the nice example presented in the triangulation paper - realities of non-exemplars is that it is very hard to get this right.

>>>>

<!-- ### Triangulation across evidence sources -->

<!-- TODO Have a look at this - potentially move to triangulation chapter -->

<!-- One key question for which multiple distinct sources of evidence were available were those looking at Laz -->

<!-- Consideration of the potential impact of the magnitude and direct of residual confounders/bias is not a major stretch from what is already happening in the assessment of the quality of evidence (GRADE) framework. Within GRADE, the overall quality of evidence can be upgraded when there is deemed to be unmeasured or residual confounding variables which reduce the. For example, if the propensity to treatment is related to comorbidity burden, but those on treatment still have better outcomes then those on control, it is likely that the true effect of the intervention is being underestimated. [@guyatt2011] -->

<!-- Without our framework and as part of the risk-of-bias assessments reported in \@ref(risk-of-bias-subheading), I attempted to records the direction of the bias, so that it could feed into the triangulation. -->

@@

### Data sources

This chapter builds on the comprehensive systematic review presented in CHapter \@re(sys-rev-heading), and incorporates the results of the analyses presented in Chapters \@ref(cprd-heading) & \@ref(ipd-heading).

### Background to the causal question

Following best practice guidance, three

Table with columns describing exposure, outcome, timepoint, other aspects (cumulative), and

This was also guideded by the forthcoming ROBINS-E tool, which has

@@

### Risk of bias tool

In addition, the tool also aim to capture the potential direction of bias for each result. Possible responses included: "Favours experimental", "Favours comparator", "Towards null", "Away from null", and "Unpredictable". Highlight that this is slightly different for the confounding domain in non-randomised studies.

These levels only apply to existing tools, not the MR tool. However, a similar approach was employed.

### Graphical representation

Where the direction of bias was unclear/could not be determined from the report, this is indicated

In order to aid with the triangulation exercise, a new method of presentation of these results was developed to enable detailed comparison across different studies contributing to the causal question.
The level of bias in each study is reported using coloured blocks, while the predicted direction of bias in that domain, categorised as towards or away from the null, is indicated using an arrow

Of note, a different approach was required for the confounding domains in the assesment tools used for non-randomised studies. Confounding in a study will either pull to the left or right, regardless of where the effect estimate is, while other domains will pull towards/away from the null (e.g. non-differential misclassification). In this case, the program accounted for the position of the effect estimate when assigning a directional arrow in relation to the positon of the effect estimate. 
For example, see Figure \@ref(), which shows an example study under the same confounding structure (protective), but with protective and harmful effect estimates. In this case, when the estimated effect is protective (Study 1), the arrow for the confounding domain indicates that the bias is pulling it away from the null. If the estimated effect is harmful, the arrow indicates that the results is being biased towards the null.

<!-- TODO Does immortal time bias not also have a consistent effect -->

@@

### Standardisation across study designs

Standardising across different measures of ex

Want best information across lipids, and so standardised across differnt measures

Fixing directions of effect - get everything pointing the same way

<!----------------------------------------------------------------------------->
### Analyses

Bias-adjusted meta-analysis [@higgins2008a]

@@

Summary of risk-of-bias/triangulation results

Percentage of domains for which a direction of bias could be assigned?

@@

Applicability/indirectness as an issue 
Compare with GRADE and cite George's example of 40% of effect predicted by MR seen when using statins for 5 years

@@

### Challenges of real-life data

Compare and contrast with the nice example presented in the triangulation paper - realities of non-exemplars is that it is very hard to get this right. ALso highlight the issue with assigning a direct of bias in many studies

### Need for new methods

Can I suggest any empirical studies that need to be performed? Average strength of immortal time bias/etc?
